### 2022-11-13 Docs/supplement.Rnw
cross ref ms and supp, model outputs

%%% start preambling . . .  %%%
\documentclass{article}

% required 
\usepackage{xr-hyper}
\usepackage{hyperref}
\externaldocument[manuscript-]{manuscript}
\usepackage{booktabs,siunitx}
\usepackage{Sweave}
\usepackage{graphicx}
\usepackage{lipsum}                     % Dummytext % https://tex.stackexchange.com/questions/9796/how-to-add-todo-notes
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}

% https://tex.stackexchange.com/questions/60209/how-to-add-an-extra-level-of-sections-with-headings-below-subsubsection
\usepackage{titlesec}
\setcounter{secnumdepth}{4}

\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

% recommended! Uncomment the below line and change the path for your computer!
\SweaveOpts{prefix.string=/Users/alina/Documents/git/localadaptclim/Docs/figure_ms/demoFig, eps=FALSE} 
%put your figures in one place! Also, note that here 'figures' is the folder and 'demoFig' is what each 
% figure produced will be titled plus its number or label (e.g., demoFig-nqpbetter.pdf')

% make your captioning look better
\usepackage[small]{caption}
\setlength{\captionmargin}{30pt}
\setlength{\abovecaptionskip}{0pt}
\setlength{\belowcaptionskip}{10pt}

% optional: muck with spacing
\topmargin -1.5cm        
\oddsidemargin 0.5cm   
\evensidemargin 0.5cm  % same as oddsidemargin but for left-hand pages
\textwidth 15.59cm
\textheight 21.94cm 
% \renewcommand{\baselinestretch}{1.5} % 1.5 lines between lines
\parindent 0pt		  % sets leading space for paragraphs
% optional: cute, fancy headers
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[LO]{November 2022}
\fancyhead[RO]{Manuscript}
% more optionals! %
%\usepackage[hyphens]{url} % this wraps my URL versus letting it spill across the page, a bad habit LaTeX has

%%% end preambling. %%%

\begin{document}
\SweaveOpts{concordance=TRUE} % For RStudio hiccups
\title{{\huge Supplements for: Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies}
\author{Ziyun Zeng \& E. M. Wolkovich}
\date{November 2022}
\maketitle

\section*{Methods}
\subsection*{Table for all studies}
\subsection*{Mapped locations for all studies}

\section*{Results}

\subsection*{Spring Lat VS. Fall Lat}
\begin{table}
\centering
\caption{Model summary for the relationship between event day of year (DOY)and provenance latitude, fitted by species, in spring (left) and fall (right).}
\todo[inline]{i'm not sure how to word this}
\begin{tabular}[t]{lcc}
\toprule
  & $DOY_{Spring}~(Latitude|Species)$ & $DOY_{Fall}~(Latitude|Species)$\\
\midrule
Intercept & \num{114.219} & \num{316.736}\\
 & {}[\num{100.511}, \num{127.680}] & {}[\num{272.545}, \num{415.373}]\\
Sigma[Species × Intercept,Intercept] & \num{1148.104} & \num{12381.073}\\
 & {}[\num{435.227}, \num{2825.532}] & {}[\num{6176.075}, \num{24554.731}]\\
Sigma[Species × Latitude,Intercept] & \num{-13.454} & \num{-307.256}\\
 & {}[\num{-41.733}, \num{-1.069}] & {}[\num{-621.049}, \num{-122.796}]\\
Sigma[Species × Latitude,Latitude] & \num{0.374} & \num{10.159}\\
 & {}[\num{0.111}, \num{1.030}] & {}[\num{5.266}, \num{33.131}]\\
\midrule
Num.Obs. & \num{671} & \num{349}\\
R2 & \num{0.903} & \num{0.961}\\
R2 Adj. & \num{0.902} & \num{0.960}\\
R2 Marg. & \num{0.000} & \num{0.000}\\
Log.Lik. & \num{-2340.362} & \num{-1217.949}\\
ELPD & \num{-2358.2} & \num{-1231.9}\\
ELPD s.e. & \num{27.8} & \num{22.1}\\
LOOIC & \num{4716.4} & \num{2463.8}\\
LOOIC s.e. & \num{55.6} & \num{44.2}\\
WAIC & \num{4716.2} & \num{2463.3}\\
RMSE & \num{7.87} & \num{8.44}\\
r2.adjusted.marginal & \num{0.902} & \num{0.960}\\
\bottomrule
\end{tabular}
\label{table:model_sf_lat}
\end{table}

\subsection*{Corrected DOY results}
\subsection*{Fitting each species \& common garden instead of just species}
\subsection*{MAT difference}
\subsection*{Might want to include how the model looked when we included studies with disagreeing provenance \& gardens}
% if we indeed include this.. need to update results section in manuscript

\end{document}